\documentclass{article}                         % Make it a report 
\usepackage{indentfirst}                       % All paragraphs indented
\usepackage{setspace}                           % For 1.5 spacing
\usepackage[hidelinks]{hyperref}      % Make table of contents clickable
\usepackage[margin=1in]{geometry} % 1 inch margin
\usepackage[english]{babel}                     % For "fancy" quotes
\usepackage[autostyle, english=american]{csquotes} % Same
\usepackage{apacite}                            % Use APA citations
\usepackage{etoolbox}                           % Misc
\usepackage{titlesec}
\usepackage{listofauthorships}
\MakeOuterQuote{"}                              % For fancy quotes

\title{Electric Vehicles in Eilat}
\author{Evan Goldstein, Valeria Kopper, Christopher Myers, Zachary Zlotnick}
\date{November 2018}

\begin{document}
\pagenumbering{gobble}
\maketitle
\newpage

\renewcommand\abstractname{Summary} % "Abstract" -> "Summary", the header lies

\pagenumbering{roman}
\tableofcontents
\newpage
\listofauthorships
\newpage
\pagenumbering{arabic}
\doublespacing

% Actual content begins here!
\section{Introduction}

Modern Israel relies on conventional (gasoline and diesel-powered) vehicles for transportation. With nearly 70\% of Israeli citizens driving to work and another 18\% taking buses, commute times have increased 30\% over the past decade \cite{Dori2018IsraeliRoads}, increasing oil consumption and worsening the country's dependence on it. The environmental impacts of oil usage are well known, but in Israel's case oil usage is more problematic. Israel produces only a few thousand barrels of oil per day and therefore imports 99\% of its oil \cite{Engber2006WhereOil}. In modern times this oil comes from Russia, Egypt, and other nearby Arabic nations -- considering Israel's troubled past with its neighbors and several oil crisis incidents and treaties \cite{Engber2006WhereOil}, Israel has serious motivation to decrease its reliance on oil in favor of more readily available sources of power.

To that end, Eilat is investigating the possibility of heavily promoting electric vehicles (EVs), along with related infrastructure and possibly a vehicle-to-grid (V2G) charging system. Briefly, electric vehicles are vehicles such as cars, buses, or even trucks that rely on traction power generated from an electric motor instead of an engine, and powered by a very high capacity lithium-ion battery instead of gasoline. Electric cars are up to 4 times more efficient than traditional internal combustion vehicles, and have been successfully demonstrated in cars such as the Tesla Roadster, with a range of up to 393 km \cite{Friel2010ManagementVehicles}. While this particular vehicle might be a higher end car, it's a good example of the viability of all-electric transportation.

Much work has already been done on investigating the benefits and drawbacks of EVs from economic, environmental, and engineering perspectives. As previously stated electric cars are well-known to be more energy efficient, but their overall impact varies with the source of the electricity. That is, if the energy is sourced from renewable sources (e.g. solar, the current best option for Eilat), all-electric is advantageous, while if the energy is sourced from fossil fuels like oil, it's more efficient to generate power on-board, such as in a hybrid electric vehicle \cite{Dincer2010EconomicOptions}, and the same goes for environmental effects. V2G charging is the idea of allowing two-way power flow between a car and the electric grid, allowing the car battery to supply power to the grid when needed, while charging when power is more readily available \cite{Mahmud2015PowerEV}. Unfortunately, the complicated equipment underpinning the idea are still under development \cite{2018Nuvve...} and have not been fully demonstrated in a complicated urban environment under real-world conditions.

The problem with applying efforts under development elsewhere to Eilat's situation is that they do not take into account Eilat's demographics and power sources. Eilat aims to transition to electric vehicles \textit{and} increase solar usage, which can interfere with part of the benefits of V2G --- charging would take place at times of decreased load, usually at night when solar panels coincidentally cannot generate power. Without adequate power management, it's entirely possible that an EV owner might wake up one morning only to find that their car has discharged completely overnight. Furthermore, without an understanding of what motivates Israelis to switch to electric cars (the average consumer is unwilling to pay extra in the name of environmental friendliness \cite{Mock2010MarketVehicles}) it will be difficult to determine how the city should properly incentivise its citizens. The applicability of V2G charging to Eilat has also not been investigated, with little to no data available on what sort of charging station frequency or range would be acceptable to those who want to switch. Finally, if Eilat wishes to switch both to solar panels and electric cars, serious upgrades to their power would be required \cite{Vardimon2011AssessmentIsrael}, but whether this is feasible or not is undetermined.

The goal of this project is to assess the technological viability and requirements of Eilat's goal to have 90\% electric vehicles by 2040, and develop a roadmap for its implementation in order to aid the city of Eilat in switching to electric vehicles. To accomplish this we aim to determine what the most effective economic incentive to switch is, to determine the maximum price that Eilat consumers would be willing to pay, what the maximum range of the cars should be, and investigate the potential for charging stations in Eilat. Although adapting the technology to Eilat's situation and convincing citizens to convert to EVs may prove challenging, it opens unique opportunities. Eilat's effort can prove V2G's potential, demonstrate a working EV-powered city, and act as a role model for other cities looking to switch to a majority of EVs.

\newpage
\section{Background}

\subsection{Transportation}[Evan G.]
Eilat is a relatively small city with a population of 47,800. Within the city center, you can walk almost everywhere. There is only one bus company, Egged. \cite{TransportationEilat} "Get Taxi", an Uber-like service, is also available. 

In March 2018, a report published by the Bank of Israel found that commute times in the country have increased by 30\% in the last decade. This is likely because private and commercial vehicles account for how 69\% of Israelis get to work nationwide. Following that, 18\% take a bus. Tel Aviv and Haifa are very close to the national numbers, but in Jerusalem, the statistic changes drastically, to 54\% car and 32\% bus \cite{Dori2018IsraeliRoads}. By comparison, in European cities like Barcelona and Brussels public transportation accounted for 40\% of all commuting and in Paris, 70\%. The report found that those who do take the bus to work do so not because it is convenient or efficient, but because they cannot afford to own a car, and therefore have no choice. The report came after the International Monetary Fund said Israel’s transportation infrastructure was insufficient and was threatening to be a drag on economic growth. 

The stop and go movements of vehicles in heavy traffic is a major case for the use of electric vehicles. A gas-powered vehicle is inefficient while idling. In the US, this accounts for 10.6 billion gallons of fuel per year, or 7.4\% of all gasoline consumption. Electric vehicles would eliminate this wasted power by using no energy while stopped. Additionally, when slowing down, EVs could recharge their batteries by converting their kinetic energy to stored electrical energy through a process called "regenerative breaking."

Israeli car ownership has increased rapidly in the past 20 years. In 1995, car ownership was climbing at 6-7\% per year --- among the fastest in the world at the time \cite{Slater1995IsraelCulture}. A 2012 report from the Israel Tax Authority showed that car owners in the center and north of the country drove the least --- around 14,000 kilometers a year on average between the Tel Aviv, Jerusalem and Haifa areas --- while car owners in far-north Metula area drove the most, at 22,000 kilometers a year. Eilat car owners, who are the farthest from the country’s center, drove the second most at 19,000 kilometers a year \cite{Schmil2012WhatHow}. The report states that between 1995 and 2011, tax revenue from cars more than doubled. Additionally, whether due to increasing green-awareness or Israel’s Green Tax Reform, cars on Israel's roads are becoming less damaging to the environment\cite{Schmil2012WhatHow}.

%Luxury car ownership has also increased rapidly, doubling in recent years \cite{Weinglass2018Rev-upIsrael}. Economist Yaron Zalicha told the Times of Israel: “The low interest rates of recent years allowed car purchasers to borrow more money and incentivized consumption over saving. …In addition, the government recently made it easier for people to import cars individually, which incentivizes the import of expensive cars” \cite{Weinglass2018Rev-upIsrael}. Some see the rise of luxury cars as being a direct result of a growing economy. As in many places, there is a stigma against owning a luxury car. It is a common stereotype in Israel that the owners of luxury cars are criminals or tax evaders. This is partly because until a new law goes into effect in 2019 it is legal in Israel to purchase apartments and cars using cash.

%Israel does not have a particularly thriving car culture scene. This is largely limited by the government. Speed limits are low, and every single modification to a vehicle, even changing the wheel size, requires special permission. New car purchases are taxed at 83\%, making the purchase of a car very expensive.

\subsection{Electric Vehicles and Solar Power}[Christopher M.]
\subsubsection{Electric Vehicles}
Electric vehicles (EVs) are an energy efficient alternative to traditional internal combustion engine vehicles (ICEVs), relying on built in lithium-ion batteries to store power and electric motors to generate traction to move the vehicle. ICEVs produce carbon dioxide (CO$^2$, a known greenhouse gas) emissions from burning gasoline, but convert 10-15\% of the chemical energy within into traction that moves the car. In comparison, a hybrid vehicle --- combining both combustion and EV technology --- can yield 30-40\% efficiency \cite{Zhu2015DistributedGrid}. For a more concrete example, gasoline has a known energy density of 9.5 kWh/liter \cite{EngineeringFactors}, and a 2018 Honda civic sedan has 46.9 liter tank with a fuel efficiency of 14.9 km/l \cite{20182018Information}. This means it can store 445.5 kWh with a range of 698 km. In comparison, a Tesla Roadster has a max range of 393 km but a battery capacity of only 60 kWh \cite{Friel2010ManagementVehicles}, making the Roadster over 4 times more energy efficient despite its lower range.

As can be seen, EVs are currently limited in both battery capacity and range, presenting a usage-awareness challenge to consumers who wish to switch. On the low end of the scale, cars can have as little as 18.8 kWh batteries with a range of only 160 km, while higher end models like the Tesla Roadster from above have 60 kWh batteries and a range of 393 km \cite{20182018Information}. For reference, Israelis use 6,893 kWh in electric power per capita as of 2018 \cite{2018Key2018}. Depending on travel habits and local availability of charging stations. The key factor in EV range is battery capacity, which unfortunately remains expensive, costing up to 1,000 USD/kWh in 2010 \cite{Mock2010MarketVehicles}, although future projections estimate that \$200-300/kWh is a realistic goal and mass production is likely to further reduce expense. Range is a known concern of drivers, who on their own have little incentive to switch to EVs \cite{Mock2010MarketVehicles}, so the best possible range and availability of charging stations is essential.

\subsubsection{Charging and Power Requirements}
EVs have significant power requirements while charging that require special accommodations. Chargers are broken down into levels 1-4, with increasing charge rate and infrastructure requirements as you progress along the scale. Modes 1 and 2 are most common currently and consist of simply plugging the EV into AC power (mode 2 has a special in-line control box for added safety) \cite{Bossche2010ElectricInfrastructure}, using hardware built into the car to convert AC to DC power. These chargers are considered less safe but are ideal for opportunity charging (plugging in your car whenever parked for an extra top-up) and don't require special infrastructure, although they can take up to 18 hours to fully charge a car at a rate of 1.4-3.3 kW \cite{Ustun2015ImpactSystems} (for reference, a typical laptop might draw 60-70 W while charging). Mode 3 requires dedicated supply equipment and is considered safer than both 1 and 2 due to built-in safety systems such as auto-poweroff when not plugged in. Mode 4 is similar but requires even heavier equipment and uses its own equipment to convert AC to DC power. Both modes 3 and 4 ("fast charging") require anywhere from 50-250 kW \cite{Ustun2015ImpactSystems, Bossche2010ElectricInfrastructure} to charge, which needs specialized infrastructure beyond current residential or industrial outlets, but at a bonus of recharging a car in as little as 10 minutes. However, this might be overkill, since even most European consumers drive 50km/day or less and Israel drives even less than Europe, so "semi-fast" charging that draws up to 22 kW is probably suitable for most people \cite{Bossche2010ElectricInfrastructure}. Regardless, the Eilat electric grid is designed to accommodate typical residential loads such as the aforementioned 6,893 kWh/year, not thousands of 22 kW loads spread across the city, so the increased energy demand and strain on existing infrastructure would be enormous.

\subsubsection{Solar Energy}
To help satisfy the enormous energy requirements of an EV powered city and reduce carbon emissions, a number of renewable energy options are available to Eilat and Israel as a whole, most notably solar power. Solar power requires large amounts of exposure to the sun and huge swathes of land to generate power effectively. Modern photovoltaic solar panels are only 10-17\% efficient \cite{Zhu2015DistributedGrid}), generating 150-200 watts per square meter, so supplying Israel with its average demand of 6.7 gigawatts would require 33-45 km$^2$ of contiguous solar panels, or widespread rooftop deployment to help satisfy a fraction of current demand \cite{Vardimon2011AssessmentIsrael}.

Unfortunately, the volatile nature of solar energy (affected by weather but entirely dependent on time of day, solar panels cannot generate power during the night) can cause problems with satisfying power demand and stabilizing power supply \cite{Lu2015IntroductionPEVs}, and the deployment of dedicated solar farms would likely require significant upgrades to the existing power grid \cite{Vardimon2011AssessmentIsrael} in both capacity and technology. Combined with the high energy demands imposed by a fleet of electric vehicles, a desperately-needed increase in solar power and 90\% EVs (as per Eilat's goal) can be expected to require significant changes to Eilat's electrical grid.

\subsubsection{Vehicle-to-Grid Charging}
One possibility for alleviating these concerns is the idea of vehicle-to-grid (V2G) charging. V2G charging centers around the idea of allowing two-way power flow between the grid and EVs connected to it for charging, instead of the traditional idea of one-way charging where you plug your car in and get a charge. Control and monitoring equipment built into the charger in collaboration with large scale grid monitoring could determine when power more readily available in the grid vs. when it's badly needed and decide whether the car charges its battery or discharges to help supplement the power supply. For instance, a well-known phenomenon is "peak time" when the work day ends, people go home, and power usage spikes. During these times, V2G-charging cars could serve as a distributed battery, discharging into the grid to relieve stress on power generators, while charging up during the nighttime in time to be driven in the morning \cite{Mahmud2015PowerEV}. While one car on its own might not have too much power, in large amounts the effect could be enormous. For reference, in 2017 Tesla built a 129 MWh battery in southern Australia which reduced service costs by 90\% \cite{Lambert2018Teslas90} --- if each EV had a battery capacity of 60 kWh, it would take exactly 2,150 of them to equal Australia's battery, or one EV for less than 5\% of Eilat's population. This makes V2G technology a serious option to mitigate a potential Eilat power crisis.

\subsection{Existing Automotive Industries and Prior Efforts}[Zachary Z.]

\subsubsection{The Existing Automotive Industry}
Israel is not traditionally considered a strong player within the global automotive industry, as they have continued to import most of their automobiles from other countries. As of 2015, 254,000 vehicles were imported to Israel - only to increase by 12\% in 2016 to a total of 286,728 vehicles \cite{}. In addition, 1,004,019 of the vehicles on Israel's roads as of 2017 were 3 years old or under (cite Halavy). This constant influx of cars due to an assumed increase in Israeli wealth has caused the few manufacturers that reside in Israel to become obsolete compared to other big manufacturers abroad including Hyundai, Toyota, and Kia (cite Car Loan World, Jerusalem Post**). Numbers like these would come as a shock to most when they consider that Israel imposes an effective tax rate of 75\% on any imported vehicle. This means that an imported \$5000 car would effectively cost the average consumer \$8850 (nbn.org). 

All of this being said, Israel does have a reputation as a breeding ground for start-ups to thrive. One example of a forward-looking start-up is "Mobileye," who has partnered with Intel and BMW to create a fully autonomous car by the year 2021. This would seemingly reduce traffic buildup in busy areas, reduce crash rates, and ensure a quicker commute (cite money.cnn). Ethical considerations would need to be taken into account however, and it would most likely only encounter widespread adoption once the cars are exponentially better drivers than humans. While the technology is not there yet, the Israeli firm has already begun testing these and it should only be a matter of time before they are available to the public.  


\subsubsection{A ``Better Place''}
With regard to EVs, there has been little effort made within Israel following the efforts made by Shai Agassi back around 2011. Agassi was a entrepreneur who had seen a future of electric vehicles in Israel - creating a start-up called "Better Place" which aimed to "eliminate range anxiety and to deliver cars cheaply" (cite broken place article **). Though these goals were attainable, the reason they failed can be attributed to the development of lofty goals along the way, and the over-confidence of CEO Shai Agassi. He was often documented to have used "Shai math," the process of taking an entirely realistic metric and making it nearly impossible to attain. Other flaws in the "Better Place" business plan lied in their allocation of investments and resources. For example, they decided to spend \$60 million on a navigation system for their vehicles, which could have been licensed to another company for \$29.95 (cite broken place article **). This is not to say that the misinformed and erratic business decisions, however it proves to be relevant when looking at the technological side of "Better Place." 

When looking at the actual implementation that Shai Agassi and the venture capitalists, it is important to note that their technological vision was sound. They wanted to create a cheap family car that had enough range for most people, and to place charging stations in sites intelligently. When a car arrived at these sites, their battery would be replaced in about 3 minutes, about the same time as a gasoline fill up, and they would be on their way again. 

%Israel is not traditionally considered a strong player within the global automotive industry; however, they are proving to be a country on the rise in this regard. As a nation, Israel is known for the environment it has provided for technology start-ups to thrive. That in mind, many global brands are deciding to invest in the "high-end automotive technology sector." Some of these global brands include Intel, looking to apply its software automation in Jerusalem; Volkswagen AG, trying to popularize the ride hailing service called "Gett,"; even BMW AG, who made a large investment in a transit app called "Moovit" \cite{Coutinho2018IsraelIndustry}. 

%With regard to the traditional facets of the automotive industry, the Israeli market also has companies specializing in Tier 1, 2, and Tier 3 supply; Tier 1 supplying companies with components for final assembly as well as aftermarket spare parts, Tier 2 supplying Tier 1 with sub-assembly components, and Tier 3 supplying simple engineered materials for Tier 2 suppliers, respectively. As of 2015, Israel has over 60 companies working in Tier 1 aftermarket supply creating spare parts and accessories such as batteries, ventilation, gears and valves. Israel also has over 50 companies that supply sub-assembly components to original equipment manufacturers (OEMs) \cite{MinistryofEconomyandIndustryStateofIsraelTheIsrael}. These companies may begin to shift over time as Israel seeks alternatives to traditional, inefficient cars, and moves toward lighter, smarter, and technologically savvy electric vehicles.

%Keeping in mind that Israel is looking to move toward smart cars and efficient vehicles, there are a number of notable trends occurring in the world right now. One shift being made is the reduction of vehicle weight by using alternative materials, given that a car that is 10\% lighter would use 6-8\% less fuel \cite{MinistryofEconomyandIndustryStateofIsraelTheIsrael}. Other popular trends lie in the information technology sector, enabling vehicles to communicate with other vehicles and become more automated. These technologies include driver assistance systems, automated and smart transportation, and information security. All of these technologies aim to reduce the inherent error that human drivers have, decreasing commute times, reducing the impact of potential crashes, and keeping personal information out of the hands of hackers \cite{MinistryofEconomyandIndustryStateofIsraelTheIsrael}. Directly related to this increase in technology within the automotive sector, there exist some ethical considerations that must be taken into account. While the idea of more efficient vehicles would not receive much opposition, technology which takes control from the driver is heavily debated. Questions regarding the culpability of the driver in certain situations come into play, as well as driver rights, manufacturer liability, and implications for future generations.

%Responding to this fundamental shift in the way the world is viewing transportation, Israel has responded with a few ideas of their own. For example, the trending navigation app "Waze" was based in Israel before they were bought out by Google in 2013. Waze encourages communication between drivers via a mobile app (not while driving obviously) where users can report traffic jams, vehicles stopped on the shoulder, hazards, and even police officers. Another Israeli company, "Cellint," monitors all active cell phones and detects their exact location on the roads every 30 seconds, collecting data to use for traffic management, research, and time-to-destination analysis. Other Israeli companies such as "Argus" aim to address the issue of information security, ensuring that data can be shared between vehicles seamlessly without interference from hackers \cite{MinistryofEconomyandIndustryStateofIsraelTheIsrael}. The list of Israel-based companies and their various applications goes on, however their goals all relate to creating safer, smoother and more efficient journeys for drivers.

%Overall, the Israeli automotive industry is a developing one. This is not to say that it is underdeveloped, but that it has focused its lens on emerging technologies and brand new global needs. International companies have focused their attention to Israel in search of new viable automobile-related technologies. In addition to the cutting-edge technologies that have sprouted in Israel, they also have made efforts to vertically integrate many of them - ensuring that there is support from initial design to final production \cite{MinistryofEconomyandIndustryStateofIsraelTheIsrael}. Israel's location enables them to get the best of both commercial worlds; they can easily trade with Asia and Europe.

\subsection{Israeli Economics}[Valeria K.]
\subsubsection{Oil Dependence}
Currently the State of Israel has a strong dependence on oil imports, which has contributed to already complicated ties with its neighboring nations in the Middle East. Israel produces incredibly small quantities of oil, as little as only a couple thousand barrels of oil a day \cite{Engber2006WhereOil}. Such a small oil production does not allow Israel to sustain more than 1\% of its oil needs, which means it imports over 99\% of its oil \cite{Engber2006WhereOil}. This generates an undesired and convoluted vulnerability for Israel. Since its establishment, the State of Israel has continuously sought oil resources closer to home in order to increase stability and safety as well as decreasing import costs. Its overall growth and increase of oil demand has further fostered Israel's need to decrease its dependence on other nations for almost the entirety of their oil resources. 

Ideally, Israel can decrease its dependence on other nations by decreasing its oil consumption. Focusing on the development of energy consumption in future years progressing towards the use of renewable energy instead of oil consumption would yield great benefits in terms of environmental, social and political aspects. This can be achieved by focusing on renewable energy production and shifting towards electric machines, such as EVs, versus oil-run machines. Israel has the opportunity to expand upon its untapped potential for renewable energy production, especially solar polar, which would grant Israel an increased independence from its oil imports and undesired ties to neighboring countries.

\subsubsection{Economic Incentives}
An important aspect to take into account in regards to implementation of EVs and V2G in Eilat is the use of economic incentives. An economic incentive can be understood to be a something that influences people in a way that causes them to behave in a certain way or make certain decisions. This is often adopted by governments in order to encourage behavior from their country’s populations, and the implementation of EVs is the perfect example for this. In order to successfully transition the city of Eilat into 90\% EV usage the city’s population’s willingness to adopt EVs is crucial. This goes to say, no matter the plan that the government develops, if the population does not choose to adopt it buy purchasing and using EVs, then there is no way to ensure the success of such a plan. 

That being said, the government has the power to incentivize people to choose EVs, and given that people generally react most strongly to matters regarding finances, economic incentives are one of the best ways to motivate people to adopt EVs. There are multiple ways of incentivizing populations to invest in EVs versus regular cars, which is ultimately the decision that determines whether Eilat can achieve 90\% EV usage by 2040. One of the most common economic incentive is tax cuts, in this case decreasing purchasing taxes or allowing for full purchasing tax exemptions (for EVs and not for traditional cars) can be incredibly effective \cite{Katsovitch2011JIMS:Highest}. Taxes on cars in Israel are among the highest in the world at 83\%, which is five times higher than most European countries. It is estimated that it can take up to 28\% of the average Israeli’s salary to be able to afford a car \cite{Katsovitch2011JIMS:Highest}, which makes this incentive even more appealing to Israeli citizens and has the potential to be especially motivating for consumers to switch to EVs. EV users can also be except road tolls. Additionally, the government could potentially allow charging stations installed around the city to be free of charge, possibly for a predetermined time period. These charging stations could also be initially free and then the consumer could increasingly (likely percentage wise) start paying for public electric charge over time. The government can also potentially offer reduced electric billing for the homes of owners of EVs. Electric charging would represent significant savings for citizens by eliminating their personal oil costs, which also serves as an important financial incentive.

Additional to economic incentives the government can also implement non-monetary incentives. A group of EV owners in Norway said the strongest incentive for them was bus lane access \cite{Bjerkan2016IncentivesNorway}. In the state of California EV users are allowed free municipal parking and access to the carpool lane \cite{TeslaElectricIncentives}. Ireland plans on implementing priority parking for EVs \cite{SustainableEnergyAuthorityofIreland2011ElectricRoadmap}.

Among these various forms of economic and non-monetary incentives there can be clear distinctions among different groups in the population. Factors such as gender, age, and education influence people in regards to what would incentivize them to purchase an EV. Identifying these and the composition of a population is crucial to determining what the most effective use of incentives for a particular city or country is. Overall if the Israeli government intends to achieve having EVs as the city's major form of transportation it must be backed up by policy (most likely monetary policy) and supportive measures (such as non-monetary incentives) for the population.

\subsection{Summary}
Overall Eilat stands much to gain by switching to electric vehicles, including decreasing its oil reliance, encouraging a reliance on Israeli-produced energy, reduced emissions, and increased political independence. However, the process of switching to 90\% electric vehicles by 2040 can be expected to be complicated. Since Eilat obviously cannot directly control its citizens purchases, the city must find a way to encourage drivers to switch to electric vehicles, using an in-depth understanding of the desired prices and vehicles ranges. In the event that the switch \textit{does} happen, Eilat can expect difficulties related to charging thousands of electric vehicles, namely problems with its power supply and electric grid. Charging cars incur a heavy electric cost, so not only must Eilat increase its solar power generation, but it must upgrade its electric grid and investigate new locations for charging stations. More research is required to develop a roadmap for Eilat's plan.

\newpage
\section{Methods}
The goal of this project is to assess the technological viability and requirements of Eilat's goal to have 90\% electric vehicles by 2040, and develop a roadmap for its implementation in order to aid the city of Eilat in switching to electric vehicles.

\subsection{Economic Incentives}[Valeria K.]
\subsubsection{Method}
In order to determine the most effective policy to incentivize the population towards transitioning to 90\% EV usage the demographics of the population of Eilat need to be broken down. To do so we plan to analyze data provided by the sponsor. By doing so we will understand the make up of the population of Eilat. Additionally, we will compare this to data regarding which population segments are the major car consumers. The overlap between these two will yield the population we want to target. Once we have this we can survey members that fall into this category in order to understand what would motivate them to purchase an EV instead of a conventional car.

\subsubsection{Analysis}
One of the most challenging and crucial aspects of developing a roadmap in order to achieve having 90\% EVs in Eilat is understanding the intricacies of the human aspect of the implementation of EVs. The technology behind it is relatively straight forward; however, people’s reactions can be much more convoluted. In order for a roadmap like this one to succeed it must address this and evaluate what is the best way to get people to embrace EVs. Focusing on this aspect through an economic scope can allow for effectively targeting the population and determining the most effective incentives available. 

Population groups determined by factors such as age, gender and education have varying reactions to incentives to adopt EVs. Determining the makeup of the population in Eilat is thus crucial to understanding what incentivizes each population segment. Utilizing various incentives targeting the city’s specific demographics will likely yield the best results. In order to do so our group will have to determine the population make up through the analysis of data provided by the project’s sponsor. This will help us understand which groups make up the majority of the population in Eilat and which groups make up the majority of car purchases. Comparing the majority of the composition to the majority of car consumers will be key to determining what are target population groups are.

Once we determine which population groups are the most relevant we plan on surveying people based on this. The purpose of the surveys is to establish what the most effective incentive is for each of these sub-populations. Given that the specific demographic determinations have not been made, the surveys will change and adapt while we are in Israel. We will focus the surveys on understanding  people's preferences. Through understanding the demographic composition of Eilat and understanding what appeals to the major car consumers we can determine which is the most effective economic incentive or combination of incentives. 

\subsubsection{Deliverables}
Through the preformed analysis and the data collected from the surveys we will be able to construct a plan of which incentives, both financial and non-monetary, are the ideal combination for Eilat to promote the incorporation of EVs. This will consist of formulating a proposal of a policy the government can implement. This can be part of a timeline that will be a part of the overall roadmap we are creating. 

\subsection{Infrastructure}[Evan G.]
Electric vehicles will require gas stations to be upgraded to EV charging stations. Transitioning from ICEVs to EVs will require an increase in electricity production to cope with the power requirements of EVs. The most effective way to increase electricity production will be to build additional solar farms. Eilat's electrical grid will require upgrades to handle the additional load of EV charging stations. 
\subsubsection{Methods}
\paragraph{Expanding solar energy production}
To determine the optimal locations for new solar farms, we will take an algorithmic approach. We will consider factors of available of land such as cost, distance from the city and power loss over transmission lines, and 

%Algorithms such as multi-objective Bees Algorithm 

\paragraph{Electrical grid updates}


\paragraph{Optimal Location for Charging stations}


\subsection{Determining Ideal Electric Car Price}[Zachary Z.]

Even if the amount of electric vehicles produced by 2040 makes up 90\% of the Eilat transportation system, that does not necessarily mean that they will all be sold. Basic supply and demand states that there is an ideal point where supply and demand meet: creating a market price for a good. While this assumes that there is little to no government intervention in the supposed market, it is a relative metric when trying to determine a fair market price of a good or service. 

\subsubsection{Method}
In order to determine the best price for electric vehicles, we need to gather data from potential consumers regarding their willingness to purchase a vehicle given a price. We will give the survey 5-10 different prices and a range of 200km, and for each price they will indicate "yes" - I would purchase an EV for that price, or "no" - I would not. We do not know yet the demographic of the participants since we are awaiting data from sponsors, and have not decided one concrete prices to try. These prices should be given without the involvement of incentives of any kind.

\subsubsection{Analysis}
Most of the work with regard to finding a fair market price falls in analyzing the data properly. This would begin with determining a percent purchase rate for each given price. Let's consider a basic example where we survey participants on prices ranging from \$26,000-30,000 (incrementing by \$1,000 yields 5 prices). If we were to take the total number of people who say "yes" for any of the 5 prices, and divide it by the total number of participants, then we can figure out the percentage of people for each price range. This would yield a sample population that we can map to the entire population of drivers in Eilat. Ideally, the more demographics we survey, and the more people in general that we survey, the better results we will get. This would also create a demand curve for a supply and demand graph.

In order to determine what the supply looks like it should be relatively straightforward. We simply need to take the total number of vehicles in Eilat, and take 90\% of that to obtain the total number of vehicles that need to be switched over to electric. This would then go on the x-axis of a supply-demand chart and would yield a supply curve. The intersection of the two curves we create would yield a fair market price for the consumer - one without incentives.

While a price without any sort of government intervention through incentive policies is better than no price, it is important to look at the effects that incentives would have on a fair price. For example, if the government were to reduce the purs

\subsubsection{Deliverables}


\subsection{Determining Acceptable Electric Car Ranges}[Christopher M]
\subsubsection{Method}
To gain an understanding of the range of electric cars that consumers expect and would be best served by, we will conduct surveys of the Eilat population. These surveys will be random samples of the Eilat population as a whole without regard to economic status or occupation as we are most interested in the opinions of the general population of Israel. However, it is well known that the range that drivers would \textit{prefer} is different than the range that drivers could reasonably live under ("range anxiety"), so the surveys will collect data on both preferred ranges and minimum acceptable ranges. This may be surveyed alongside questions of electric car price in order to determine whether consumers would be willing to accept a lower range car in exchange for reduced cost.

More specifically, the surveys will ask questions about Israeli driving habits (e.g. "How many kilometers do you drive per week on average?") first, move on to asking about long trips and frequency thereof, then suggest a list of possible ranges (0-75 km, 75-150 km, etc.) and ask what they need at a minimum and what they would prefer. The first two subjects are not intended for genuine data gathering but instead will encourage the participant to reflect on their actual usage of their vehicle. Combined with a brief overview of electric car costs and their ranges (low end to high end models), this should help participants right range anxiety and give more conservative estimates. Care should be taken to ensure that price ranges shown are not too intimidating to the average consumer; that is to say, there is no need to show the price of a high end luxury electric car if only a fraction of the population could afford it.

\subsubsection{Analysis}
Data gathered from these surveys will most importantly consist of information on the minimum acceptable ranges for Eilat citizens and what their preferred range would be, with the secondary bonus of up-to-date information on Eilat citizens' driving habits. Averaging these values and calculating basic statistics information (median, standard deviation, etc.). The gap between minimum and preferred range is also of use, as it serves as a good indicator of Eilat citizen's vulnerability to range anxiety. For instance, if minimum range is on average a third of what preferred range is, we could anticipate heavy resistance to rollout of early (modern) EVs on the basis that most citizens would find their ranges too limiting. If, on the other hand, we find that the two are very close, it indicates little to no range anxiety and the question of EV rollout becomes less a question of technology delays and more a question of consumer economics. Considering that batteries and other EV technology remains under constant development, this could prove a serious boost to Eilat's goals.

\subsubsection{Deliverables}
With up-to-date information on Eilat citizens' travel habits and more importantly their minimum/preferred ranges, we can construct a more effective timeline as part of our roadmap for Eilat's 2040 goal. This information can be combined with Eilat demographics and known EV ranges and prices to determine when it's most likely that Eilat citizens could start switching to EVs. For instance, if the average Eilat citizen wants to be able to drive at least 150 km in one trip but affordable EVs that can manage that aren't expected until 2030, we could set a definite point on our timeline. This point in time may prove to be a turning point in Eilat's plan, so depending on the results of our analysis, we may recommend preparation leading up to this point and action beginning immediately at it.

\subsection{Conclusion}


% Bibliography -- automatically managed in APA style, on a new page
\newpage
\bigskip
\bibliographystyle{apacite}
\bibliography{refs}

\end{document}
